Computational design of RNA-based oscillatory circuits

### 2015-08-24 MScWriteup.tex
writing up model section

\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes}

\usepackage[caption=false]{subfig}

@@

% The paper headers
\markboth{hi}%
{Optimisation of Car Electronics}

@@

@@

The report is structured as follows. In section \ref{The sRNA regulatory system} we review the sRNA regulatory system we will consider. In section \ref{Results and Discussion} 
we propose a set of ODE's to model the system, and estimate its unknown parameters by fitting to time series data. Finally, in section \ref{Conclusions and Further work}, we conclude, and suggest directions for further work.

\section{The sRNA regulatory system}
\label{The sRNA regulatory system}
In bacteria, one mechanism by which gene expression is regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA. This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA. In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS. Translation may be self repressed by this `tail' of the mRNA folding over and binding across the RBS, forming a stem loop in the mRNA and preventing the Ribosome from binding (Fig. \ref{RBS}). This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind (Fig. \ref{RBS}). In summary, the presence of the sRNA positively regulates gene expresssion.

@@

\begin{figure*}[b!]
\centering
\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
\caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS. In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$. These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E. coli} discussed. These chemical repressors are themselves disabled by two chemicals, aTC and IPTG. In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production. We see a 'double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA). Thus presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.  Image reproduced from \cite{Rodrigo2012}.}
\label{ANDGate}
\end{figure*}

%%%
%NEW SECTION%
%%%
\section{Results and Discussion}
\label{Results and Discussion}

@@

\subsection{Model Derivation}

 \begin{itemize}
 \item introduce the proposed ode model, explain its biology.
 \item Biological background, taken from those 3 papers.
 \end{itemize}

>>>>

\usepackage{float,amsfonts,multicol,fancyvrb,lastpage,ragged2e,url,color,lipsum,enumerate,todonotes,chemfig,fixltx2e}
\usepackage[caption=false]{subfig}

@@

% The paper headers
\markboth{hi}%
{Computational design of RNA-based oscillatory circuits}

@@

%%%
%%%
%%%
%%%
%%%
%NEW SECTION%
%%%
%%%
%%%
%%%

@@

The report is structured as follows. In the remainder of this section we review the sRNA regulatory system we will consider. In section \ref{Results and Discussion} we propose a set of ODE's to model the system, and estimate its unknown parameters by fitting to time series data. Finally, in section \ref{Conclusions and Further work}, we conclude, and suggest directions for further work.

\subsection{The sRNA regulatory system}
\label{The sRNA regulatory system}
In bacteria, one mechanism by which gene expression is regulated is as follows \cite{Soper2010}: In order for a bacterial mRNA to be translated into a protein, the Ribosome must initially bind to the mRNA (Fig. \ref{RBS}). This occurs at the Ribosome Binding Site (RBS) \cite{Shine1974}, a specific nucleotide sequence found on the mRNA. In an mRNA there is an untranslated region of nucleotides at the 5' end of the molecule (the UTR), upstream of the RBS. Translation may be self repressed by this `tail' of the mRNA folding over and binding across the RBS, forming a stem loop in the mRNA and preventing the Ribosome from binding (Fig. \ref{RBS}). This self repression may be released with an sRNA which binds to the same region on the mRNA - the new conformation of the sRNA:mRNA complex uncovers the RBS, allowing the Ribosome to bind. In summary, the presence of the sRNA positively regulates gene expresssion.

\begin{figure*}[t]
\centering
\includegraphics[trim = 30 600 30 0,page=4,clip = true]{pnas1203831109.pdf}
\caption{A logical AND gate formed from a self repressed mRNA, and an sRNA which uncovers its RBS. In this system, transcription of the sRNA (transRNA) and mRNA (5'-UTR,GFP) are controlled by two promoter regions, $P_{LtetO-1}$ and $P_{LlacO-1}$. These are disabled by the presence of two chemical repressors, TetR and LacI, found naturally in the strain of \textit{E. coli} discussed. These chemical repressors are themselves disabled by two chemicals, aTC and IPTG. In the notation of the diagram, a barred line indicates repression, and an arrowed line indicates production. We see a 'double negative' in aTc repressing TetR, which itself represses transcription of the sRNA (likewise for IPTG and the mRNA). Thus presence of the sRNA and mRNA are controlled by the presence of aTc and IPTG, which can be experimentally introduced to the cell.  Image reproduced from \cite{Rodrigo2012}.}
\label{ANDGate}
\end{figure*}

@@

Although a qualitative understanding of this system exists \cite{Rodrigo2012}, it is of interest to attempt a quantitative understanding of the genetic circuit involved. Such an understanding would allow, for example, tailoring of the system in reponse to design requirements, by altering the values of the important parameters of the model. By changing which sRNA-mRNA device is used in the system, it would also allow exploration of the relationship between the thermodynamic properties of each device, and the model's rate constants.

%%%
%%%
%%%
%%%
%%%
%NEW SECTION%
%%%
%%%
%%%
%%%

\section{Results and Discussion}
\label{Results and Discussion}

@@

\subsection{Model Derivation}
We propose a simple model, consisting of a set of ODE's with mass action kinetics \cite{UriAlon}, to describe the system. Tables \ref{ModelParameters},\ref{StateVariables} give complete descriptions of parameter and state variable meanings.

\begin{align}
\frac{ds}{dt} &= \frac{N\alpha_{T}}{f_{T}} y(t)-(\mu + \delta_{s})s -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m \label{eq:s}\\
\frac{dm}{dt} &=  \frac{N\alpha_{L}}{f_{L}}x(t)-(\mu + \delta_{m})m -k_{\mathrm{on}}sm +k_{\mathrm{off}}s:m  \label{eq:m}\\
\frac{ds:m}{dt} & = k_{\mathrm{on}}sm  - (k_{\mathrm{off}}+ k_{\mathrm{hyb}})s:m  -(\mu + \delta_{sm} )s:m \label{eq:sm}\\
\frac{dc}{dt} & = k_{\mathrm{hyb}}s:m  -(\mu + \delta_{c})c  \label{eq:c} \\
\frac{dp}{dt} & = \beta m +f_{s}\beta c -(\gamma + \mu + \delta_{g})p - \frac{v_{z}p}{K_{z}+p+g}  \label{eq:p} \\
\frac{dg}{dt} & = \gamma p - (\mu + \delta_{g})g - \frac{v_{z}g}{K_{z}+p+g} \label{eq:g} \\
z &= z_{0} +\frac{g}{\theta} \label{eq:z}
\end{align}
  Based on the reaction mechanism in Fig. \ref{reactionscheme}, the hybridization of the sRNA and mRNA first into an unstable complex, then a stable one, is modelled in eqs.\ref{eq:s} - \ref{eq:c}. The initial binding is modelled as a reversible reaction with forward and backward rates $k_{on}$ and $k_{off}$.
\begin{align*}
sRNA \chemsign+ mRNA\chemrel[$k_{on}$][$k_{off}$]{<>} sRNA:mRNA_{unstable} 
\end{align*}\\
\noindent
after which the stabilization is modelled as an irreversible reaction with rate $k_{hyb}$. 
 \begin{align*}
sRNA:mRNA_{unstable} \chemrel[$k_{hyb}$]{->} sRNA:mRNA_{stable} 
\end{align*}

 In addition, these complexes are given degradation rates, $\delta_{s}$, $\delta_{m}$, $\delta_{sm}$, $\delta_{c}$, and dilution of chemical concentrations due to cell growth are modelled with a dilution rate $\mu$.

 Control of the system by aTc and IPTG are modelled by $y(t)$ and $x(t)$ in eqs.\ref{eq:s}, \ref{eq:m}. $y(t)$ models the response of the sRNA transcription rate to a time varying aTc concentration - it is normalised to lie between $1$ and $f_{T}$, and is typically sigmoid in response to aTc concentration \cite{Rodrigo2012}. $x(t)$ behaves analogously for the mRNA.

\begin{table}[h]
\renewcommand{\arraystretch}{1.3}
\caption{Model Parameters}
\label{ModelParameters}
\centering
\begin{tabular}{| l | l | l |}
\hline Parameter &  Units & Definition  \\
\hline \hline N & & plasmid copy number  \\
\hline $z_{0}$ &  AFU & baseline fluorescence  \\
\hline $\alpha_{L}$ & nM/min & Max translation rate of Pllac01 promoter  \\
\hline $\alpha_{T}$  &  nM/min  & Max translation rate of Pltet01 promoter \\
\hline $f_{L}$ &  & Max llacI fold  \\
\hline $f_{T}$ &  & lltet fold\\
\hline $\delta_{g}$  & /min  & gfp degradation rate  \\
\hline $\gamma$ &  /min & gfp maturation rate  \\
\hline $v_{z}$ & nM/min & degradation constant of clpx  \\
\hline $K_{z}$   &   nM/min & Dissociation constant of clpx  \\
\hline $\theta$  &   nM/AFU & ratio between gfp amount and fluorensence  \\
\hline $\mu$ &  /min & dilution rate  \\
\hline $\delta_{m}$ &  /min & mrna degredation rate  \\
\hline $\delta_{s}$ &  /min & srna degredation rate  \\
\hline $\delta_{sm}$ &  /min & unstable s:mrna degredation rate  \\
\hline $\delta_{c}$ &  /min & stable s;m degredation rate  \\
\hline $k_{on}$ &   /min & s+m $\rightarrow$ s:m rate \\
\hline $k_{off}$ &  /min & s+m $\leftarrow$ s:m rate \\
\hline $k_{hyb}$ &  /min & s:m $\rightarrow$ c rate \\
\hline $\beta$ &   /min & basal translation rate \\
\hline $f_{s}$ & & srna repression fold \\
\hline
\end{tabular}
\end{table}

\begin{table}
\renewcommand{\arraystretch}{1.3}
\caption{State Variables}
\label{StateVariables}
\centering
\begin{tabular}{| l | l |}
\hline State variable &  Definition  \\
\hline\hline $s$ &  Definition  \\
\hline $m$ &  Definition  \\
\hline $s:m$ &  Definition  \\
\hline $c$ &  Definition  \\
\hline $p$ &  Definition  \\
\hline $g$ &  Definition  \\
\hline $z$ &  Definition  \\
\hline $y(t)$ &  Definition  \\
\hline $x(t)$ &  Definition  \\
\hline
\end{tabular}
\end{table}
